Leaf traits drive differences in biomass partitioning among major plant functional types

### 2014-12-15 manuscript.Rnw
Removed figure1b, figure1c, figureS5
Reshuffle of figures (removed Fig 3, replaced with S7); updated var part tables.

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
% \usepackage[T1]{fontenc}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lineno}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\bibliographystyle{plain}

@@

\item Here we use BAAD, large database blah blah, three main PFTs, many biomes, wide range in plant size, good coverage of climate space where woody plants occur (Fig 1a and S1). We included several key variables from BAAD (Fig 1b).

@@

\item We analyzed climate effects in two ways, biome, quantitative variables (MAP,MAT,MI,mgdd0)
\item We did not find differences between biomes in $M_F/A_S$ (Fig S5 - now S7!), but some small inconsistent differences in $A_F/A_S$ (corresponding to slightly lower LMA in tropical vs. temperate evergreen angiosperms). SHOULD THIS BE FIGURE 4? Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height (Table S7 SHOULD BE IN MAIN TEXT). 

@@

\item We found that PFTs differed in LMF, and this was proportional to leaf mass per area, a leading trait within the leaf economics spectrum. Indeed, if LMA was used instead of PFT, we were able to explain more variation in $M_F/A_S$ and $A_F/A_S$ (Table S7). It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. 
\item We found very modest effects of climate on biomass partitioning, somewhat inconsistent with Reich et al 2014. It is consistent with the idea that LMA is more flexible than biomass partitioning (Poorter). We argue that climate directly or indirectly affects LMA, which then leads to differences in LMF, because LAR is less variable. Plants seem to target a certain total leaf area display per unit aboveground biomass investment, but we do stress that there is still much variation unaccounted for (Figure S histograms), and that other studies certainly have uncovered small-scale effects of climate on leaf area to sapwood area (pine studies etc), an important component of $A_F/A_S$. Our results hold at the PFT level, which is relevant for DGVMs.
\item What consequences for DGVMs? There is a multitude of allocation models (de Kauwe). Together with assumptions on leaf and stem turnover rates, this gives estimates of biomass partitioning (S2 methods). So far it is common to choose between different models largely based on philosophical arguments (Franklin 2012). We argue that relationships such as in Fig 2a should be used to benchmark and constrain the various allocation models. The data we use here is freely available (Falster), which should aid in the models being tested, with the hope that they perhaps converge.

@@

% Fig 3 to SI.
% Fig S7 to main text

@@

\section{Tables}

@@

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure3.pdf}
    \caption{Plant functional types diverge strongly in mass-based partitioning, but converge in area-based partitioning. Shown are histograms of leaf mass per area (an index of leaf construction cost) ($M_F/A_F$), leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/M_S$) grouped by the three major plant functional types. Arrows indicate means (same as in Fig.~\ref{fig:figure2}).}
    \label{fig:figure3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure4b.pdf}
    \caption{Modest effects of climate on three size-invariant variables, compared to the very large effect of plant functional type on mass-based variables (leaf mass per area, LMA, and $M_F/A_S$), and convergence in area-based variables ($A_F/A_S$). See SI Methods for calculation of MI and growing season temperature. Solid lines are generalized additive model fits (with a base dimension of 3, larger values resulted in biologically implausible relationships). Grey areas are approximate 95\% confidence intervals, calculated by the gam after averaging each variable by species/dataset combination (but all raw individual plant data are shown in the plots).}
  \label{fig:figure4}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS1.pdf}
    \caption{Global coverage of the climate space by the dataset. Methods are as in Fig 1 in main text, symbols are colored by vegetation type as reported by the data contributors.}
\label{fig:figureS1}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS5.pdf}
    \caption{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel). Each symbol is an individual plant. Lines are generalized additive model fits.}
  \label{fig:figureS5}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS6.pdf}
    \caption{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel). Individual plant sata were averaged by species/study (the same species occuring in multiple datasets thus appears multiple times). Lines are linear regressions with 95\% confidence intervals. Both regressions were significant, but R2 was 0.28 in the left panel, and only 0.02 in the right panel. The coefficient of variation between species was 0.19 for log-transformed MF/AS, and only 0.099 for AF/AS.}
  \label{fig:figureS6}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS7.pdf}
    \caption{Mean leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/A_S$) by PFT and major biome. No leaf area estimates were available for boreal deciduous angiosperms. DA = deciduous angiosperm, EA = evergreen angiosperm, EG = evergreen gymnosperm, Trop. = tropical, Temp. = temperate. Error bars are 95\% confidence intervals for the mean, estimated from a linear mixed-effects model using species within study as the random effect. Letters denote significant differences (at $\alpha = 0.05$)}
    \label{fig:figureS7}
\end{figure}

@@

@@

<<results='asis'>>=
print(xtable(table_pipemodel_varpart, label="varpartpipe",
             caption="Explained variance in the ratio of leaf mass to basal stem area, and leaf area to basal stem area. The marginal R2 is that explained by the fixed effects only (H = plant height, B = biome, PFT = plant functional type. The conditional R2 is that explained by fixed and random effects (species within study). All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."),
      include.rownames=FALSE, caption.placement="top")
@

<<results='asis'>>=
print(xtable(table_LMFLAR_varpart, label="varpartlmflar",
      caption="Explained variance in the leaf mass fraction (leaf mass / aboveground biomass) and leaf area ratio (leaf area / aboveground biomass). See Table~\\ref{varpartpipe} for more explanation. For plant height, we also added the squared term as a fixed effect, and the interaction with PFT or LMA, because the relationship was non-linear."),
      include.rownames=FALSE, caption.placement="top")
@

>>>>

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lineno}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\bibliographystyle{plain}

@@

\item Here we use BAAD, large database blah blah, three main PFTs, many biomes, wide range in plant size, good coverage of climate space where woody plants occur (Fig 1a and S1). We included several key variables from BAAD (Fig 1b). Crucially, the data are for individuals, which is more relevant to DGVMs than plot-level estimates, because ...

@@

\item We analyzed climate effects in two ways, biome, quantitative variables (MAP,MAT), and by biome (tropical / temperate/ boreal), which is a simple classification taking into account both MAT and MAP.
\item Regardless of how we analyzed it, we found only weak climate effects. We did not find differences between biomes in $M_F/A_S$ (Fig 3), but some small inconsistent differences in $A_F/A_S$ (corresponding to slightly lower LMA in tropical vs. temperate evergreen angiosperms). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height (Table S7 SHOULD BE IN MAIN TEXT). 

@@

\item We found that PFTs differed in LMF, and this was proportional to leaf mass per area, a leading trait within the leaf economics spectrum. Indeed, if LMA was used instead of PFT, we were able to explain more variation in $M_F/A_S$ and $A_F/A_S$ (Table S7). It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. 
\item We found very modest effects of climate on biomass partitioning, somewhat inconsistent with Reich et al 2014. It is consistent with the idea that LMA is more flexible than biomass partitioning (Poorter). We argue that climate directly or indirectly affects LMA, which then leads to differences in LMF, because LAR is less variable. Plants seem to target a certain total leaf area display per unit aboveground biomass investment, but we do stress that there is still much variation unaccounted for (Figure S histograms), and that other studies certainly have uncovered small-scale effects of climate on leaf area to sapwood area (pine studies etc), an important component of $A_F/A_S$. Our results hold at the PFT level, which is relevant for DGVMs.
\item What consequences for DGVMs? There is a multitude of allocation models (de Kauwe). Together with assumptions on leaf and stem turnover rates, this gives estimates of biomass partitioning (S2 methods). So far it is common to choose between different models largely based on philosophical arguments (Franklin 2012). We argue that relationships such as in Fig 2a should be used to benchmark and constrain the various allocation models. The data we use here is freely available (Falster), which should aid in the models being tested, with the hope that they perhaps converge.

@@

@@

\clearpage
\section{Tables}

@@

<<results='asis'>>=
print(xtable(table_varpart2, label="varpartpipe",
             caption="Explained variance in the ratio of leaf mass to basal stem area, and leaf area to basal stem area. The R2 shown is that explained by the fixed effects only (H = plant height, B = biome, PFT = plant functional type.  All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."),
      include.rownames=FALSE, caption.placement="top")
@

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure3.pdf}
    \caption{Mean leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/A_S$) by PFT and major biome. No leaf area estimates were available for boreal deciduous angiosperms. DA = deciduous angiosperm, EA = evergreen angiosperm, EG = evergreen gymnosperm, Trop. = tropical, Temp. = temperate. Error bars are 95\% confidence intervals for the mean, estimated from a linear mixed-effects model using species within study as the random effect. Letters denote significant differences (at $\alpha = 0.05$)}
    \label{fig:figure3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure4.pdf}
    \caption{Modest effects of climate on three size-invariant variables, compared to the very large effect of plant functional type on mass-based variables (leaf mass per area, LMA, and $M_F/A_S$), and convergence in area-based variables ($A_F/A_S$). See SI Methods for calculation of MI and growing season temperature. Solid lines are generalized additive model fits (with a base dimension of 3, larger values resulted in biologically implausible relationships). Grey areas are approximate 95\% confidence intervals, calculated by the gam after averaging each variable by species/dataset combination (but all raw individual plant data are shown in the plots).}
  \label{fig:figure4}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS1.pdf}
    \caption{Global coverage of the climate space by the dataset. Methods are as in Fig. 1 in main text, symbols are colored by vegetation type as reported by the data contributors.}
\label{fig:figureS1}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS5.pdf}
    \caption{Plant functional types diverge strongly in mass-based partitioning, but converge in area-based partitioning. Shown are histograms of leaf mass per area (an index of leaf construction cost) ($M_F/A_F$), leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/M_S$) grouped by the three major plant functional types. Arrows indicate means (same as in Fig.~\ref{fig:figure2}).}
    \label{fig:figures5}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS6.pdf}
    \caption{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel). Individual plant data were averaged by species/study (the same species occuring in multiple datasets thus appears multiple times). Lines are linear regressions with 95\% confidence intervals. Both regressions were significant, but $R^2$ was 0.28 in the left panel, and only 0.02 in the right panel.}
  \label{fig:figureS6}
\end{figure}

@@

\clearpage

@@
